Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for

### 2020-08-27 Scripts/Initial_Analysis.Rmd
continue sst emerg. analysis
Working on the multilinealmodel

#### Project's Library
packages <- c(
  "dbemImport",
  "tidyverse",
  "here", # for dbem_import `here()`
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  "geosphere", # estimate distances between points `distm()`
  "ggrepel",
  "zoo", # for average mean
  "parallel", # for mclapply,
  "sf", # for mapping
  "st", # for mapping
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  "zeallot", # for Juanito's map
  "gmt", # for estimating distances between points
  "viridis",
  "cowplot"
)

@@

EEZ_CellID <- my_path("G", extra_path = "Spatial/V_Lam", name="EEZ_CellID.xlsx", read = TRUE) %>%
  clean_names()
colnames(EEZ_CellID) <- c("eezid","index")

# SAU name relatioal table (Fish For Visa)
sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="SAU_matching_names.csv", read = TRUE)

@@

ModelRCP <- "GFDL85"

getEnvVar=function(yr,ModelRCP,ensemble,Variable){

  filepath <- paste("Z:/JULIANO_NEYMAR/PristineSeasData/Climate",ModelRCP,"",sep="/")

@@

# Get taxon list
Taxon_List_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))

# clean names for function
Taxon_List_emer <- gsub("_.*","",Taxon_List_emer)

@@

@@

# Clean names to match data
Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T)

# Species depth clasification

Spp_depth <-fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
  clean_names()

@@

# Load all sp in one table
Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
  filter(
    n_ens == 10
  ) %>% 
  left_join(Neighbours,
            by = c("eez_name","eez_neighbour")
  ) %>% 
  rename(sf_eez_name=eez_name) %>% 
  left_join(Matching_names,
            by=  "sf_eez_name") %>% 
  clean_names()

@@

# Per EEZs
Eez_d_hist <- Eme_Res %>% 
  group_by(eezid= eez_id ,sau_fishing_entity,tresh) %>% 
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(
    mean_yr = mean(emerging_yr,na.rm=T),
    sd_yr = sd(emerging_yr,na.rm=T),
    n_taxa = length(unique(taxon_key))
    ) %>% 
  gather("metric","value",mean_yr:n_taxa) %>% 
  filter(metric != "n_taxa") %>% 
  mutate(
    reverse = abs(value-2100)  + 2000
  )

@@

>>>>

#### Project's Library
packages <- c(
  "dbemImport",
  "tidyverse",
  "here", # for dbem_import `here()`
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  "geosphere", # estimate distances between points `distm()`
  "ggrepel",
  "zoo", # for average mean
  "parallel", # for mclapply,
  "sf", # for mapping
  "st", # for mapping
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  "zeallot", # for Juanito's map
  "gmt", # for estimating distances between points
  "viridis",
  "cowplot",
  "rfishbase"
)

@@

EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
  clean_names()
colnames(EEZ_CellID) <- c("eezid","index")

# SAU name relatioal table (Fish For Visa)
sau_clean_names <- my_path("G", extra_path = "Spatial/SAU/", name="matching_names.csv", read = TRUE)

@@

getEnvVar=function(yr,ensemble,Variable){

  filepath <- paste("Z:/DATA/Environmental data/GFDL_CMIP5_Large_Ensemble/ens",ensemble,"",sep="/")

@@

# Get taxon list
Taxon_list_emer <- list.files(my_path("R",extra_path = "Emergence_2005"))

# clean names for function
Taxon_list_emer <- gsub("_.*","",Taxon_list_emer)

# SAU exploited species list
Exploited_spp_list <- my_path("G","SAU","exploited_species_list.csv",read = T) %>%
  clean_names() %>% 
  pull(taxon_name)

# Get habitat preference from fishbase
Spp_env_pref <- rbind(species(Exploited_spp_list, #species from rfishbase package
                              fields = species_fields$habitat[4])
                      ) %>% 
  mutate(taxon_name = Exploited_spp_list) %>% 
  left_join(my_path("G","SAU","exploited_species_list.csv", read = T) %>%
              clean_names(),
            by = "taxon_name")

@@

# Clean names to match data
Matching_names <- my_path("G","Spatial/SAU", "matching_names.csv", read = T) %>% 
  rename(eezid = eez_id)

Fishing_names <- Matching_names %>% 
  group_by(sau_fishing_entity,fishing_entity_id) %>% 
  summarise(n=n()) %>% 
  select(-n)

@@

# eez_centroids
Eez_centroids <- my_path("D", extra_path = "Spatial/", name="eez_centroids.csv", read = TRUE) %>% 
  rename(sf_eez_name = eez_name) %>% 
  left_join(Matching_names) %>% 
  select(eezid,latitude,longitude)

# Species depth clasification (SAU-fishbase)
Spp_depth <-fread(paste(my_path("D"),"SAU/SppTaxonName.csv", sep ="")) %>% 
  clean_names()

# Time of Emergence of SST using methods here
Toe_sst <-fread(paste(my_path("D"),"Climate/ToE_sst.csv", sep ="")) %>% 
  clean_names()

# Biogeographic regions (from Rey 2020)
Biogeo_reg <-fread(paste(my_path("D"),"Climate/SPATIAL_PARTITION_OCEAN.csv", sep ="")) %>% 
  clean_names() %>% 
  rowid_to_column("index")

# Get eez to index relational data
EEZ_CellID <- my_path("G", extra_path = "Spatial/DBEM", name="EEZ_CellID.xlsx", read = TRUE) %>%
  clean_names()
colnames(EEZ_CellID) <- c("eezid","index")

Eez_biogeo <- EEZ_CellID %>% 
  left_join(Biogeo_reg,
       by = "index") %>% 
  # Deal with NaNs in data
  mutate(biome = ifelse(biome == "NaN",0,biome)) %>% 
  filter(biome > 0) %>% 
  # Get frequency of biome per index
  group_by(eezid,basin,biome) %>% 
  summarise(nbiom=n(),
            nbasin=n()) %>% 
  # Select the more frequent biome per eez
  group_by(eezid,basin) %>% 
  top_n(1,nbiom) %>% 
  group_by(eezid,biome) %>% 
  top_n(1,nbasin) %>% 
  ungroup() %>% 
  select(-nbiom,-nbasin)

# Species catch trend (from FishForVisa)

Catch_trend <- my_path("D","Species","Catch_trend.csv", read = T) %>% 
  clean_names() %>% 
  filter(taxon_key %in% Taxon_list_emer) %>% 
  rename(sau_fishing_entity = territory) %>% 
  left_join(fishing_names,
            by = "sau_fishing_entity") %>% 
  filter(status != "No Trend") %>% 
  mutate(category = ifelse(status %in% c("Developing","Rebuilding"),"Category A",
                         ifelse(status == "Max Exploited","Category B",
                                ifelse(status %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
                         )
         )
  )

@@

# Load all sp in one table
Eme_Res <- bind_rows(lapply(Paths, fread)) %>% 
  filter(
    n_ens == 10
  ) %>% 
  left_join(Neighbours,
            by = c("eez_name","eez_neighbour")
  ) %>% 
  rename(sf_eez_name=eez_name) %>% 
  left_join(Matching_names,
            by=  "sf_eez_name") %>% 
  clean_names() %>% 
  rename(eezid = eez_id)

@@

# Per EEZs
Eez_d_hist <- Eme_Res %>% 
  group_by(eezid ,sau_fishing_entity,tresh) %>% 
  # summarise(n_tax = length(unique(taxon_key))) %>% 
  summarise(
    mean_yr = mean(emerging_yr,na.rm=T),
    sd_yr = sd(emerging_yr,na.rm=T),
    n_taxa = length(unique(taxon_key))
    ) %>% 
  gather("metric","value",mean_yr:n_taxa) %>% 
  filter(metric != "n_taxa") %>% 
  mutate(
    reverse = abs(value-2100)  + 2000
  )

@@

### Can we explain the ToE?

```{r multiple_reg, eval = F, echo = F}

# Possible explanatory variables

## Habitat association (DemersPelag)
## Dem Vs Pelagic (Dem_pel)
## SST time of Emer (Toe_sst)
## Stock Status [FishForVisa] (Catch_trend)
## Biogeographic location (Biogeo_reg)

# AIC value; model between predictive capacity and complexity (i.e., number of explanatory variables) is given by the model with lowest AIC value

# -------------------------------- #
#### Prepare data
# -------------------------------- #

Lm_data <- Eme_Res %>% 
  filter(tresh == "Tresh_one") %>% 
  left_join(Spp_depth) %>% 
  select(taxon_key,tresh,sf_eez_name,sau_fishing_entity,fishing_entity_id,emerging_yr,eezid,dem_pel) %>% 
  left_join(Toe_sst,
            by = c("eezid","tresh")
            ) %>% 
  left_join(Catch_trend,
            by = c("taxon_key","sau_fishing_entity","fishing_entity_id")
            ) %>% 
  left_join(Eez_biogeo,
            by = "eezid") %>% 
  left_join(Spp_env_pref,
            by = "taxon_key") %>% 
  left_join(Eez_centroids,
            by = "eezid") %>% 
  mutate(latitude = abs(latitude)) %>% 
  filter(tresh == "Tresh_one") %>% 
  distinct()

# -------------------------------- #
#### Build models
# -------------------------------- #

# Model 1; all variables
M_all_var <- lm(emerging_yr ~ sst_toe + status + biome + dem_pel + DemersPelag, data = Lm_data)

all_sum <- summary(M_all_var)
all_aic <- AIC(M_all_var)
all_r <-  all_sum$adj.r.squared

# Model 2; non demPel
M2 <- lm(emerging_yr ~ sst_toe + status + biome + dem_pel, data = Lm_data)

M2_sum <- summary(M2)
M2_aic <- AIC(M2)
M2_r <-  M2_sum$adj.r.squared

# Model 3; non biome
M3 <- lm(emerging_yr ~ sst_toe + status + dem_pel + DemersPelag, data = Lm_data)

M3_sum <- summary(M3)
M3_aic <- AIC(M3)
M3_r <-  M3_sum$adj.r.squared

# Model 4; non status
M4 <- lm(emerging_yr ~ sst_toe + biome + dem_pel + DemersPelag, data = Lm_data)

M4_sum <- summary(M4)
M4_aic <- AIC(M4)
M4_r <-  M4_sum$adj.r.squared

# Model 5; non sst
M5 <- lm(emerging_yr ~ status + biome + dem_pel + DemersPelag, data = Lm_data)

M5_sum <- summary(M5)
M5_aic <- AIC(M5)
M5_r <-  M5_sum$adj.r.squared

# Model 6; all + basin
M6 <- lm(emerging_yr ~ sst_toe + status + biome + basin + dem_pel + DemersPelag, data = Lm_data)

M6_sum <- summary(M6)
M6_aic <- AIC(M6)
M6_r <-  M6_sum$adj.r.squared

# Model 7; Categories instead of status
M7 <- lm(emerging_yr ~ sst_toe + category + biome + dem_pel + DemersPelag, data = Lm_data)

M7_sum <- summary(M7)
M7_aic <- AIC(M7)
M7_r <-  M7_sum$adj.r.squared

# Model 8; All variables + ccentroid lat
M8 <- lm(emerging_yr ~ sst_toe + category + biome + dem_pel + DemersPelag + latitude, data = Lm_data)

M8_sum <- summary(M8)
M8_aic <- AIC(M8)
M8_r <-  M8_sum$adj.r.squared

tibble(
  "model" = c("all variables", "no demPel","no biome","no status","no sst","all + basin","Categories","lat centroids"),
  "AIC" = c(all_aic,M2_aic,M3_aic,M4_aic,M5_aic,M6_aic,M7_aic,M8_aic),
  "rsquared" = c(all_r,M2_r,M3_r,M4_r,M5_r,M6_r,M7_r,M8_r)
)

20282659
